### 2020-01-17 paper.tex
added stuff about connections betwee forms and tori

\documentclass[a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{textcomp}
%\usepackage[dutch]{babel}
\usepackage{amsmath, amssymb, amsthm}
%\usepackage{geometry}
\usepackage{tikz-cd}
\usepackage{mdframed}
\usepackage{microtype}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{todonotes}
% figure support
\usepackage{import}
\usepackage{xifthen}
\usepackage{pdfpages}
\usepackage{transparent}
\newcommand{\incfig}[1]{%
	\def\svgwidth{\columnwidth}
	\import{./figures/}{#1.pdf_tex}
}

\pdfsuppresswarningpagegroup=1

\newtheoremstyle{theoremdd}% name of the style to be used
  {7pt}% measure of space to leave above the theorem. E.g.: 3pt
  {7pt}% measure of space to leave below the theorem. E.g.: 3pt
  {\itshape}% name of font to use in the body of the theorem
  {0pt}% measure of space to indent
  {\bfseries}% name of head font
  {}% punctuation between head and body
  { }% space after theorem head; " " = normal interword spae
  {\thmname{#1}\thmnumber{ #2}\thmnote{ (#3)}}

\newtheoremstyle{definitiondd}% name of the style to be used
  {7pt}% measure of space to leave above the theorem. E.g.: 3pt
  {7pt}% measure of space to leave below the theorem. E.g.: 3pt
  {\rmfamily}% name of font to use in the body of the theorem
  {0pt}% measure of space to indenlt
  {\bfseries}% name of head font
  {}% punctuation between head and body
  { }% space after theorem head; " " = normal interword space
  {\thmname{#1}\thmnumber{ #2}\thmnote{ (#3)}}

  \newtheoremstyle{remarkdd}% name of the style to be used
  {7pt}% measure of space to leave above the theorem. E.g.: 3pt
  {7pt}% measure of space to leave below the theorem. E.g.: 3pt
  {\rmfamily}% name of font to use in the body of the theorem
  {0pt}% measure of space to indent
  {\itshape}% name of head font
  {}% punctuation between head and body
  { }% space after theorem head; " " = normal interword space
  {\thmname{#1}\thmnumber{ #2} \thmnote{ (#3)}}

\theoremstyle{theoremdd}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{property}[theorem]{Property}
\theoremstyle{definitiondd}
\newtheorem{definition}[theorem]{Definition}
\theoremstyle{remarkdd}
\newtheorem{example}[theorem]{Example}
\newtheorem{remark}[theorem]{Remark}

\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\ltr}{\par \noindent \framebox[1\width]{ $\implies$ } \hspace{.2cm}}
\newcommand{\rtl}{\par \noindent \framebox[1\width]{ $\impliedby$ } \hspace{.2cm} }
\newcommand{\exercise}[1]{\begin{mdframed}
	\textit{Exercise:} #1
\end{mdframed}}
\newcommand{\bigset}[2]{ \left\{ #1 \;\middle|\; #2 \right\} }

\DeclareMathOperator{\LC}{LC}
\DeclareMathOperator{\FC}{FC}
\DeclareMathOperator{\coef}{Coef}
\DeclareMathOperator{\coker}{coker}
\DeclareMathOperator{\Id}{Id}
\DeclareMathOperator{\Ann}{Ann}
\DeclareMathOperator{\im}{Im}
\DeclareMathOperator{\SL}{SL}

\author{Micha\"el Maex}
\date{\today}
\title{Modular Curves}
\begin{document}

\maketitle

\bigskip

\tableofcontents

\pagebreak

\section{Introduction}

This is the report of the journey of a student learning about elliptic curves. 
This report should be thought of as my notes while working my way through the first chapters of \emph{A First Course in Modular Forms} by Fred Diamond and Jerry Shurman. 
This book is my main reference. 
Most of this report is paraphrasing this book, but I've added some details where I thought the book was lacking. 
I've also tried to add some visual intuitions I developed to aid my understanding of this material. I hope potential readers will find these useful as well.

Modular forms, elliptic curves and modular forms are three seemingly unrelated concepts that have surprising connections. 
The goal of this report is to understand these objects and how they are related.

\section{Modular Forms}
Modular forms are function on the complex half plane $\mathcal{H} $ that are invariant in some sense under the action of a group (or subgroups of) called the modular group.

\subsection{The Modular Group and its Action}
We begin by introducing the modular group and its action on $\mathcal{H} \cup \{\infty\} $.

\begin{definition}
	The \emph{Modular group} are the $2\times 2$ matrices with integer coefficients and determinant $1$.
	 \[
		 \SL_2(\Z) =   \bigset{\begin{pmatrix} a& b\\ c& d \end{pmatrix}}{a, b, c, d \in \Z, ac - bd = 1}
	.\] 
\end{definition}
It can be shown that the modular group is generated by \[
	\begin{pmatrix}  1 & 1 \\ 0 & 1 \end{pmatrix} 
	\text{ and }
	\begin{pmatrix} 0 & -1 \\ 1 & 0 \end{pmatrix} 
.\] 
For a proof of this we refer to the appendices. 
We now define how this group acts on the complex half plane. 
\begin{definition}
	Let  $\gamma = \begin{pmatrix} a & b \\ c & d \end{pmatrix} $ be any element of $\SL_2\Z$ and $\tau$ be any element of $\mathcal{H} \cup \{\infty\} $.
	The multiplication of these two elements is defined as 
	\[
		\gamma(\tau) = \gamma \cdot \tau = \frac{a \tau + b}{ c \tau + d}
	.\] 
\end{definition}
Note that this group action is well defined. First we have to check that this every $\tau \in \mathcal{H} \cup \{\infty\}  $ gets mapped to another element of $\tau \in \mathcal{H} \cup \{ \infty\}$. 
Note that is is sufficient to check this for the action of the generators, which induce the maps \[
\tau \mapsto  \tau+1 \text{ and } \tau \mapsto \frac{-1}{\tau}
\]
from which it is obvious.
We can also easily see that $I_2 \cdot \tau = \tau$. 
We also see that for $\gamma = \begin{pmatrix} a & b \\ c & d \end{pmatrix} $ and $\beta = \begin{pmatrix} f & g \\ h & i \end{pmatrix} $ is holds that 
\begin{align*}
	\gamma \left( \beta \cdot  \tau\right) &=  \gamma\cdot \frac{f\tau + g}{h \tau + i} \\
					       &= \frac{a\left( \frac{f \tau + g}{h\tau + i} \right) + b}{c \left( \frac{f \tau + g}{h \tau + i} \right) + d} \\	
					       &= \frac{(af + bh)\tau + ga + bi}{(cf + dh) \tau + cg + di} \\
					       &= \begin{pmatrix} af + bh & ag + bi \\ cf + dh & cg + di \end{pmatrix} \cdot \tau \\
					       &= (\gamma \beta)\cdot \tau \\
.\end{align*}

\subsection{Subgroups of the Modular Group}
Quite often it will be useful to restrict our focus to subgroups of the modular group. 
The type of subgroup we'll most interested in are the so-called \emph{congruence subgroups}.
To define these we first have to define the so called \emph{principal congruence subgroups}.  
\begin{definition}
	For any $N \in \N$ we define the \emph{principal congruence subgroup of level $N$} to be \[
		\Gamma(N) = \bigset{\begin{pmatrix} a & b \\ c & d \end{pmatrix} \in \SL_2(\Z)}{\begin{pmatrix} a & b \\ c & d \end{pmatrix}  \equiv \begin{pmatrix} 1 & 0 \\ 0 & 1 \end{pmatrix}  (\mathrm{mod}\; N)}
	.\] 
\end{definition}

Further we define two subgroups
\begin{definition}
	For every $N \in \N$ we define $\Gamma_0$ and $\Gamma_1$ to be 
	\begin{align*}
		\Gamma_0(N) &= \bigset{\begin{pmatrix} a & b \\ c & d \end{pmatrix} \in \SL_2(\Z)}{\begin{pmatrix} a & b \\ c & d \end{pmatrix}  \equiv \begin{pmatrix} * & * \\ 0 & * \end{pmatrix}  (\mathrm{mod}\; N)} \\
		\Gamma_1(N) &= \bigset{\begin{pmatrix} a & b \\ c & d \end{pmatrix} \in \SL_2(\Z)}{\begin{pmatrix} a & b \\ c & d \end{pmatrix}  \equiv \begin{pmatrix} 1 & * \\ 0 & 1 \end{pmatrix}  (\mathrm{mod}\; N)}
	.\end{align*}
\end{definition} 
Note that 
\[
	\Gamma(N) \subset \Gamma_1(N) \subset \Gamma_0(N) \subset  \SL_2\Z
.\] 
We state without proof that $[\Gamma_1(N): \Gamma(N)] = N$, $[\Gamma_0 (N): \Gamma_1(N)]= \phi(N)$, $[\SL_2\Z: \Gamma_0(N)] = N \prod_{p \mathbin | \N}(1 + 1 /p)$. 
From this it follows that $[\Gamma(N): \SL_2\Z]= N^3 \prod_{p \mathbin | N} 1 - 1 /p^2$.
We will now give the general definition of \emph{congruence subgroups}. 
 \begin{definition}
	 A subgroup $\Gamma \subset \SL_2\Z$ is a \emph{congruence subgroup} if $\Gamma(N) \subset \Gamma$ for some $N \in \N$. In this case we say that $\Gamma$ is a congruence subgroup of level $N$. 
\end{definition}
\subsection{Defining Modular Forms}\label{sec:defining_modular_forms}

Some functions on $\mathcal{H} $ are very well behaved under this action.

\begin{definition}
	Let $k$ be an integer. 
	A meromorphic function $f: \mathcal{H}  \to \C$ is \emph{weakly modular of weight $k$ } if \begin{equation} \label{eq:weakly_modular}
		f(\gamma(\tau)) = (ct + d)^{k}f(\tau)
	,\end{equation}
	for all $\gamma = \begin{pmatrix} a & b \\ c & d \end{pmatrix}  \in \SL_2(\Z)$ and $\tau \in \mathcal{H} $. 
\end{definition}

It's clearly enough so check \eqref{eq:weakly_modular} only for the generators of $\SL_2\Z$. I.e.\begin{equation}\label{eq:sufficient_weakly_modular}
	f(\tau + 1) = f(\tau) \text{ and } f\left(-\frac{1}{\tau}\right)  = \tau^{k}f(\tau)
.\end{equation} 
From this we see that weakly modular functions are $\Z$ periodic.

We will now introduce the weight-$k$ operator on functions $[\gamma]_k$.
 \begin{definition}
	 Let $f: \mathcal{H}  \to \C$ be a function,  $\gamma = \begin{pmatrix} a & b \\ c & d \end{pmatrix} \in \SL_2\Z$ and $k \in \Z$ then we define the function $f[\gamma]_k$ to be 
	  \begin{align*}
		  f[\gamma]_k: \mathcal{H}  &\longrightarrow \C \\
		  \tau &\longmapsto (c\tau + d)^{-k}f(\gamma(\tau))
	 .\end{align*}
\end{definition}
This allows us to characterize  \emph{weakly modular of weight  $k$} as follows.
 \begin{lemma}
	 A function $f: \mathcal{H}  \to \C$ is weakly modular of weight $k$ if and only if for every $\gamma \in \SL_2$: \[
		 f[\gamma]_k = f
	 .\] 
\end{lemma}
In light of this characterization we extend to definition of weakly modular to congruence subsets of $\SL_2\Z$.

\begin{definition}
	Let $\Gamma$ be a congruence subset of $\SL_2\Z$. We say that a function $f: \mathcal{H}  \to \C$ is \emph{weakly modular of weight $k$ with respect to $\Gamma$} if for every  $ \gamma \in \Gamma$:
	 \[
		 f[\gamma]_k = ff[\gamma]_k 
	.\] 
\end{definition}

We are now ready to give the definition of a modular form. 
\begin{definition}\label{def:modular_form}
	Let $\Gamma$ be congruence subgroup of $\SL_2\Z$
	A function $f: \mathcal{H}  \to \C$ is a \emph{modular form of weight $ k$ with respect to $\Gamma$} if 
	\begin{enumerate}
		\item $f$ is holomorphic on $\mathcal{H} $,
		\item $f$ is weakly modular of weight $k$ with respect to $\Gamma$, 
		\item $f[\gamma]_k$ is holomorphic at $\infty$ for every $\gamma \in \SL_2\Z$. 
	\end{enumerate}
	We denote the set of modular froms of weight $k$ with respect to $\Gamma$ as $\mathcal{M} _k(\SL_2\Z)$.

	If $\Gamma = \SL_2\Z$ we say simply that $f$ is a modular form of weight $k$. 
\end{definition}
The stament 'holomorhic at $\infty$' deserves some explanation. There are many equivalent ways of defining it, but for our purposes we will take this to mean that $f$ allows a fourier expansion (not too suprising as  $f$ is $\Z$ periodic)\footnote{The real picture behing holomopich at $\infty$ is that, due to $f$ being $\Z$-periodic, $H$ can be 'rolled up' on the unit disk, where $\infty$ gets mapped to the center. In this picture  $f$ being holomorphic at $\infty$ means that  $f$ allows an extension which holomorphic on the center of the disk.}, i.e.\ \[
	f(\tau) = \sum_{n=0}^{\infty} a_n e^{n 2\pi i \tau}
,\]
for some $a_n \in \C$.
Another equivalent statement is that the limit $\lim_{\im(\tau) \to \infty} f(\tau)$ exists and is finite.

Note that it is only seccesary to verify the 3rd point for a representitive of every coset $\alpha \Gamma$ as by the weak modularity  $f[\alpha]_k = f[\alpha \gamma]_k$ for every  $\gamma \in \Gamma$. 
If  $f$ is a modular form (so $\Gamma = \SL_2\Z$) then we just have to verify that $f$ is holomorphic at $\infty$.
Note that $\mathcal{M}(\Gamma)$ is a $\C$-vector space. 
It turns out that the last condition of \cref{def:modular_form} makes $\mathcal{M} (\SL_2\Z)$ of finite dimension.  
\begin{lemma}\label{lem:product_modular_forms}
	Let $f$ be a modular form of weight $k$ and $g$ a modular form of weight $l$, then 
	$fg$ is a modular form of weight  $k + g$.
\end{lemma}
\begin{proof}
	The three conditions from definition \ref{def:modular_form} must be verified. 
	\begin{enumerate}
		\item The product of holomorphic functions is holomorphic. So $ fg$ is holomorphic on $\mathcal{H} $. 
		\item Take any  $\gamma = \begin{pmatrix} a & b \\ c & d \end{pmatrix} \in \SL_2\Z$ and $\tau \in \mathcal{H} $. Then 
			\begin{align*}
				(fg)(\gamma \cdot \tau) &= f(\gamma \cdot \tau) g(\gamma \cdot \tau)   \\
							&= (c\tau + d)^{k} f(\tau) (c \tau)^{l} g(\tau) \\
							&= (c\tau + d)^{k+l} (fg)(\tau) \\
			.\end{align*}
			So $fg$ is weakly modular of weight $k+l$. 
		\item We know that $\lim_{\im(\tau) \to \infty} f(\tau)$, $\lim_{\im(\tau) \to \infty} g(\tau)$ exists and is finite. Hence $\lim_{\im(\tau) \to \infty} fg (\tau)$ exists and is finite.		
	\end{enumerate}
\end{proof}

Finally we introduce a special type of modular called a cusp. The names comes from connection to elliptic curves \todo{check that is are elliptic curves and not modular curves} that we will discuss later.
\begin{definition}\label{def:cusp}
	A \emph{cusp form} is a modular form $f$ (with respect to a congruence subgroup  $\Gamma$) such that for every  $\alpha \in \SL_2\Z$  the leading coefficient in its Fourier expansion is 0. 
	 The $\C$-vector space of cusp forms of weight $k$ is denoted $\mathcal{S} _k(\Gamma)$.
\end{definition}
Again, it is sufficient to check this for a representitive for every coset $\alpha \Gamma$.

\subsection{Examples of Modular Forms}
We may wonder whether these types of functions exist. One trivial modular form of weight $k$ is the zero function
, but the question remain whether there are non trivial modular forms of a given weight.
If $k$ is odd the answer is negative as $f(\tau) = f(-I_2 \tau) = (-1)^{k} f(\tau) = 0$. 
If we know a few modular forms of small weight, we can construct modular forms of higher weights using \cref{lem:product_modular_forms}.

However, we can construct a non trivial modular form for every even weight greater than 2.
\begin{definition}
	The \emph{Eisenstein series of weight $k$} ($k$ even and  $k >2$) is the function
	\begin{align*}
		G_k: \mathcal{H} &\longrightarrow \C \\
		\tau &\longmapsto \sum_{\substack{(c,d) \in \Z^2 \\ (c,d)\ne (0,0) }} \frac{1}{(c\tau + d)^{k}}
	.\end{align*}
\end{definition}
We of course have to check that this is indeed a modular form of weight $k$. 
Basic analysis can be used to show that the summation is absolutely convergent for every  $\tau \in \mathcal{H}$ and uniformly convergent on compact sets.
It follows that $G_k$ is holomorphic.
Recall from \cref{sec:defining_modular_forms} that it is suficient to check weak modularity for  $\gamma = \begin{pmatrix} 1 & 1 \\ 0 & 1 \end{pmatrix} $ and $\gamma = \begin{pmatrix} 0 & -1 \\ 1 & 0  \end{pmatrix} $, i.e.\ we have to show that \cref{eq:sufficient_weakly_modular} holds.
We see 
\begin{align*}
	G_k(\tau + 1) &= \sum_{\substack{(c, d) \in \Z^2 \\ (c, d) \ne (0,0)}} \frac{1}{(c (\tau + 1) + d)^{k}}\\
			 &= \sum_{\substack{(c, d) \in \Z^2 \\ (c, d) \ne (0,0)}} \frac{1}{(c \tau + (d+c))^{k}}\ \\
.\end{align*}
But this is just a permutation of the terms of the previous sequence, as $\Z^2 \to \Z^2: (c,d) \mapsto (c, c+ d)$ is a isomorphism of groups.
Hence $G_k(\tau + 1) = G_k(\tau)$.
We have to check the second equation as well.
\begin{align*}
	G_k\left(\frac{-1}{\tau}\right) &=  \sum_{\substack{(c,d) \in \Z^2 \\ (c,d) \ne 0}} \frac{1}{(-c / \tau + d)^{k}}\\
			     &= \tau^{k} \sum_{\substack{(c, d) \in \Z^2 \\ (c,d)}} \frac{1}{(d \tau - c)^{k}} \\
			     &= \tau^{k}G_k(\tau) \\
\end{align*}
The last equality is again true as we can permute the terms as the summation is absolutely convergent.

A lengthy calculation that we will omit shows that \begin{equation}\label{eq:expansion_eisenstein}
	G_k(\tau) = 2 \zeta(k) + 2 \frac{(2\pi i)^{k}}{(k-1)!}\sum_{n = 1}^{\infty} \sigma_{k-1}(n)  e^{n 2 \pi i \tau}
,\end{equation}
whenever $ k>2$ and $k$ is even, where $\zeta$ is the Riemann zeta function and $\sigma$ is the sum of divisors function, defined as
\[
	\sigma_{k}(n) = \sum_{\substack{d \mathbin | n \\ d > 0}} d^{k}
.\] 
We've yet to find an example of a cusp. We can do this by taking the right linear combination of two independent elements of $\mathcal{M} _k(\SL_2\Z)$ in a way such that the first coefficient cancels.
In this spirit we define the \emph{discriminant function}
\begin{align*}
	\Delta = (60 G_4)^3 - 27(140G_6)^2
.\end{align*}
Note that this a modular form of weight 12. 
Using \cref{eq:expansion_eisenstein} we see that the first two coefficients are $0, (2 \pi)^{12}$. 
Hence it is a non-trivial cusp of weight 12.

\section{Lattices and Complex Elliptic Curves}
The main idea of this section is to construct tori by considering $\C$ as an additive group modulo a certain kind of embedding of $\Z^2$. This embedding of $\Z^2$ is called a lattice.  
\begin{definition}
	A \emph{lattice in $\C$} is additive subgroup $\Lambda = \omega_1 \Z \oplus \omega_2 \Z$ where $\omega_1, \omega_2 \in \C$ are linearly independent over $\R$ and $\omega_1 / \omega 2  \in \mathbb{H}$
\end{definition}
\begin{definition}
	A \emph{complex torus} is a quotient of the complex plane by a lattice, \[
		\C / \Lambda = \bigset{z + \Lambda}{z \in \C} 
	.\] 
\end{definition}
This object is called as torus because of a geometric intuition. 
If one considers the parallelogram spanned by $\vec{\omega_1}, \vec{\omega_2}$ we see that the opposing edges are `glued' together. 
So topologically this space is homeomorphic to the torus.

\begin{figure}[h]
	\centering
	\incfig{lattice_complex_plane}
\end{figure}

The modular group tells us when two two lattices are equal. 
\begin{lemma}\label{lem:equality_of_lattices}
	Let $\Lambda = \omega_1\Z \oplus \omega_2\Z$ and $\Lambda' = \omega_1'\Z \oplus \omega_2\Z$ be two lattices. 
	Then $\Lambda = \Lambda'$ if and only if there exists an $\begin{pmatrix} a & b \\ c& d \end{pmatrix} \in \SL_2\Z$ such that \[
	\begin{pmatrix} \omega_1' \\ \omega_2' \end{pmatrix}  = 
	\begin{pmatrix} a & b \\ c& d \end{pmatrix} 
	\begin{pmatrix} \omega_1 \\ \omega_2 \end{pmatrix} 
	.\] 
\end{lemma}
\begin{proof}[Proof. (exercise 1.3.1)]
	\ltr
	We know that $\omega_1' \in \Lambda$, there there are $a, b \in \Z$ such that $\omega_1' = a \omega_1 + b \omega_2$. Similarly there are $c, d \in \Z$ such that $\omega_2' = c \omega_1 + d \omega_2$. 
	So  \[
	\begin{pmatrix} \omega_1' \\ \omega_2' \end{pmatrix}  = 
	\begin{pmatrix} a & b \\ c& d \end{pmatrix} 
	\begin{pmatrix} \omega_1 \\ \omega_2 \end{pmatrix} 
	.\] 
	We still have to argue that the matrix is in $\SL_2\Z$, i.e. has determinant 1. 
	Analogously we can find $f, g, h, i \in \Z$ such that \[
	\begin{pmatrix} \omega_1  \\ \omega_2 \end{pmatrix}  = 
	\begin{pmatrix} f & g \\ h& i \end{pmatrix} 
	\begin{pmatrix} \omega_1' \\ \omega_2' \end{pmatrix} 
	.\]
	Hence \[
	\begin{pmatrix} \omega_1 \\ \omega_2 \end{pmatrix}  = 
	\begin{pmatrix} f & g \\ h & i \end{pmatrix} 
	\begin{pmatrix} a & b \\ c & d  \end{pmatrix} 
	\begin{pmatrix} \omega_1 \\ \omega_2 \end{pmatrix} 
	.\]  
	From this it follows that the determinants must be either $1$ or $-1$. 
	We can see that is must be $1$ as the imaginary part of $\omega_1 / \omega_2$ and $\omega_1' / \omega_2'$ must be both positive.
	\rtl
	We can use the fact that the lattices are groups. We know that the $\Lambda'$ is generated by $\omega_1'$ and $\omega_2'$.
	We further know that $\omega_1' = a \omega_1 + b \omega_2$ so $\omega_1' \in \Lambda$. 
	Similarly we see that $\omega_2' \in \Lambda$.
	So $\Lambda' \subset  \Lambda$.

	By inverting the matrix we using completely analogous reasing that $\omega_1 \in \Lambda'$ and $\omega_2 \in \Lambda'$. So $\Lambda \subset \Lambda'$.
\end{proof}

A complex torus is a Riemann surface. This a not a report about Riemann surfaces, but to not completely shove everything under the rug I will quickly explain what a Riemann Surface is.
Recall that that smooth manifolds are spaces that are locally homeomorphic to $\R^{n}$ and equipped with a smoothly compatible atlas. 
In the same spirit Riemann surfaces are objects that look locally like $\C$ and equipped with an atlas that is holomorphicly compatible.
The complex tori are Riemann surfaces.

I will mention one theorem about Riemann surfaces without proof. 
\begin{theorem}
	Let $f$ be a holomorphic map between compact Riemann surfaces. Then $f$ is either a map to a single point or a surjection.
\end{theorem}

\subsection{Morphisms between Complex Tori}

In our study of complex tori, holomorphic maps between them are of great importance. 
In light of this we want to know how such a maps look like. It turns our there linear transformations with some restrictions. 
\begin{proposition}
	Let $\phi: \C \ \Lambda \to \C \ \Lambda'$ be a holomorphic map between complex tori. 
	Then $\phi(z + \Lambda) = mz + b + \Lambda'$ for some $m, b \in \C$ such that $m \Lambda \subset  \Lambda'$. 
	The map is invertible if and only if $m \Lambda = \Lambda'$.
\end{proposition}
\begin{proof}
	The idea is to use topology to lift the map. As $\R^2 \simeq \C$ is the universal covering space of the torus we know that we can lift $\phi$ to a holomorphic map $\tilde \phi: \C / \Lambda \to \C$. We then can make this into a homolomorphism by first projecting $\C$ onto $\C / \Lambda$. We get a holomorphic map  $\tilde \phi: \C \to \C$.
	Take any $\lambda \in \Lambda$. 
	Then the map $f_\lambda(z) = \tilde\phi(z + \lambda) - \tilde \phi(z)$. We know that $f_{\lambda}$ is continous, but we also know that is must map to an element of $\Lambda'$, which is discrete, so it must be constant. 
	We can differentiate to find that $\tilde\phi'(z+ \lambda) = \tilde\phi'(z)$ for every $\lambda \in \Lambda$, hence it is  $\Lambda$ periodic.

	This makes $\tilde\phi'$ bounded so its constant. 
	So $\tilde \phi(z) = mz + b$ for some $m, b \in \C$. 
	Recall that $f_{\lambda}$ is is constant and equal to element of $\Lambda'$. 
	$\phi(z+\lambda) - \phi(z) = m\lambda \in \Lambda'$. 
	Hence $m \Lambda \subset \Lambda'$.

	We will now prove the second part of the theorem.
	\ltr Suppose that the inclusion  $m\Lambda \subset \Lambda'$ is proper. 
	Take a $\lambda \in \Lambda' \setminus m\Lambda$. 
	So $\lambda / m \not\in  \Lambda$. But $\phi(\lambda m) = b + \Lambda' = \phi(\lambda)$. Thus $\phi$ is not injective. So there is no inverse.
	\rtl It can be easily checked that the map $\psi: \C / \Lambda' \to \Lambda: z + \Lambda \mapsto  \frac{z-b}{m} + \Lambda$ is an inverse. 
\end{proof}

\subsection{Modular Forms and Complex Tori} \label{sec:modular_forms_and_complex_tori}
We now have a notion of \emph{morphisms} between complex tori. We can now ask ourselves the question how equivalence classes of isomorphic complex tori look like.
It turns out that every complex torus $\C / \Lambda$ is \emph{isormorphic} to a torus of the form $\C / \Lambda_\tau$, where $\Lambda_\tau = \tau \Z \oplus \Z$, for some $\tau \in  \mathcal{H}$. 
Indeed, by definition $\Lambda = \omega_1 \Z \oplus \omega_2 \Z$ for some $\omega_1, \omega_2 \in \C$ such that $\omega_1 / \omega_2 \in \mathcal{H} $. Let $\tau = \omega_1 / \omega_2$. 
Then $\frac{1}{\omega_1} \Lambda = \Lambda_\tau$. 
So the map  \[
\phi: \frac{\C}{ \Lambda} \to \frac{\C}{\Lambda_\tau}: z + \Lambda \mapsto \frac{z}{\omega_1} + \Lambda_\tau
\]
is an isomorphism.

This $\tau$ however may not be unique. Suppose that for some $\tau' \in \mathcal{H} $ the torus $\C / \Lambda_{\tau'}$ is also isomorphic to $\C / \Lambda$. In that cause $\C / \Lambda_{\tau'}$ is also isomorphic to $\C / \Lambda_{\tau}$. 
Then for some $m$ it must be the case that $m \Lambda_{\tau'} = m\tau' \Z \oplus m \Z = \Lambda_\tau$. 
By \cref{lem:equality_of_lattices} this implies that there is some $\gamma = \begin{pmatrix} a & b \\ c& d \end{pmatrix}  \in \SL_2\Z$ \[
\begin{pmatrix} m\tau' \\ m \end{pmatrix}  = \begin{pmatrix} a & b \\ c& d \end{pmatrix}  \begin{pmatrix} \tau \\ 1 \end{pmatrix} 
.\] 
So \[
\tau' = \frac{m \tau'}{m} = \frac{a \tau + b}{c \tau + d} = \gamma\cdot \tau
.\] 
Hence we see that $\tau$ and $\tau'$ lie in the same orbit of the action of $\SL_2\Z$. 
Converesely $\C / \Lambda_\tau$ and $\C / \Lambda_{\gamma\cdot \tau}$ will always be isomorhic.
Hence we conclude that there is a bijctice correspondence of isomorphism classes of complex tori and orbits of $\SL_2\Z$ acting on $\mathcal{H} $.\[
\begin{tikzcd}
	\left\{ \substack{\text{isormorphism classes} \\ \text{of complex tori}} \right\} \arrow[r, leftrightarrow] & 
	\mathcal{H} / \SL_2\ 
\end{tikzcd}
.\] 
\subsection{Isogenies}
If we only consider the maps where the contsant term $b$ is $0$ then these maps are clearly group morphisms between $\C / \Lambda$ and  $\C / \Lambda'$. 
As a result, the a holomorphic group morphism between these two tori exists if and only if $m\Lambda = \Lambda'$ for some $m \in \C$. 
A nonzero holomorphic group morphism between complex tori is called an \emph{isogeny}.

\begin{lemma}
	Isogenies are surjective and have a finite kernel. 
\end{lemma}
\begin{proof}
	Let $\phi: \C / \Lambda \to \C / \Lambda'$ be an isogeny. 
	From the previous discussion we know that $\phi(z+\Lambda) = mz + \Lambda'$ for some $m \in\C$ such that $m \Lambda \subset  \Lambda'$. 
	For every  $x + \lambda'$ it's clear that  $\phi( x / m + \Lambda) = x + \Lambda'$. 
	So  $\phi$ is surjective. 
	Suppose that $\phi(z + \Lambda) = 0 + \Lambda'$. Then  $m z \in \Lambda'$. So  $z \in  \Lambda'/m$. 
	We also know that  $\Lambda \subset \Lambda' / m$. 
	The inclusion map is an injective group morphism from $\Z^2$ to  $\Z^2$ which must be of finite index. 
	So there are only finitely many $z + \Lambda$ such that  $\phi(z + \Lambda) = 0$.
\end{proof}
To gain a more intuitive understanding of isogenies, it is helpful to think of them as functions on the plane that map lattices \emph{into} lattices.
In this context, we see that an isogeneny can scale and rotate the lattice but we cannot skew it as illustrated in \cref{fig:isogeny}.
\begin{figure}[h]
	\centering
\incfig{isogeny_complex_plane}
\caption{An isogeny from a lattice $\Lambda$ generated by $\omega_1$ and $\omega_2$ to a lattice $\Lambda'$ generated by $\omega_1'$ and $\omega_2'$}.
\label{fig:isogeny}
\end{figure}

\begin{example}
	The so called \emph{multiply by integer maps} are isogenies. These are maps of the form \[
		[N]: \C / \Lambda \to \C / \Lambda: z + \Lambda \mapsto  Nz + \Lambda
	,\] where $N$ is an integer. 
	Note that the kernel of this map is $\frac{1}{N}\Lambda / \Lambda$ and is isomorphic to $\Z / N \Z \times  \Z / N \Z$. We denote this kernel as $E[N]$ and is the set of all $N$ torsion points in $\C / \Lambda$. This can be seen in \cref{fig:Ntorsion_subgroup}.
\end{example}
	\begin{figure}[h]
		\incfig{NTorsion_subgroup}	
		\caption{The $N$-torsion subgroup $E[N]$ of a complex torus, generated by  $\omega_1 / N$ and $\omega_2 / N$.}
		\label{fig:Ntorsion_subgroup}
	\end{figure}
\begin{example}
	Another example are the \emph{Cyclic quotient maps}. 
	Let  $\C / \Lambda$ be a complex torus and  $N \in \N$.
	If take an $N$ torsion point, $\lambda$, of $\C / \Lambda$ (a point in $\frac{1}{N} \Lambda / \Lambda$) then we can consider the subgroup $C$ generated by $\lambda$. Note that $C + \Lambda$ is a super lattice of $\Lambda$. Then the map \[
		\pi: \C \ \Lambda \to \C / (C + \lambda): z + \Lambda \to z + C + \Lambda
	\]  
	is an isogeny with kernel $C$. 
\end{example}

These two examples are very important, as they allow the construction of every isogeny. 
\begin{theorem}
	Let $\phi: \C / \Lambda \to \C / \Lambda': z + \Lambda \mapsto mz + \Lambda'$ be an isogeny. Then $\phi = \psi \circ \pi \circ [n]$, where  $\psi$ is an isomorphism, $\pi$ is a cyclic quotient map and $[n]$ is a multiply by integer map.
\end{theorem}
\begin{proof}
	Denote $K = \ker \phi$. We know that $K = m^{-1} \Lambda' / \Lambda$.
	Let $N$ be the order of $K$. So we know that every point of $K$ is $N$ torsion. 
	Hence \[
		K \subset N^{-1} \Lambda / \Lambda \simeq \Z / N \Z \oplus \Z / N \Z. 
	\]
	By the structure theorem for finite abelian groups we know that  \[
	K \cong \Z / n \Z \times \Z / n n' \Z
	\]
	for some $n, n'\in \N$. 
	This means that $nK \cong \Z / n' \Z$. Denote $\pi: \C / \Lambda \to \C / (nK + \Lambda)$ the cyclic quotient map. Now define the map \[
		\psi: \frac{\C}{nK + \Lambda} \to \frac{\C}{\Lambda'}: z + nK \mapsto  \left(\frac{m}{n}\right) z + \left( \frac{m}{n} \right)(nK + \lambda)
	,\] 
	which is a well defined isomorphism as $mK + m\Lambda = \Lambda'$. 
	I now claim that $\phi = \psi \circ \pi \circ [n]$. Indeed, take any $z + \Lambda$ then \[
		z + \Lambda \mapsto nz + \Lambda \mapsto  nz + nK + \Lambda \mapsto  mz + \Lambda' \phi(z + \Lambda)
	.\]  
\end{proof}
\todo{dual isogenies, weil pairings}
\section{Ellpitic curves and their connection to Complex Tori}
An elliptic curve is a cubic curve of the form \[
y^2 = 4x^3 - g_2x - g_3
,\] 
where $g_2, g_3 \in \Z$ and $g_2^3 - 27 g_3^2 \ne 0$. 
These objects are related to complex tori, by the meromorphic functions from the complex tori to the complex plane. 
A mermorphic function $\phi: \C / \Lambda \to \C \cup \{\infty\} $ are naturally identified with a  $\Lambda$-periodic function $\phi: \C \to \C \cup \{\infty\}$.
One may wonder how a non-trivial example of such a function looks like. 
One example of particular interest is the \emph{Weierstrass $\wp$-function}. 
For a given lattice  $\Lambda$ we define this function to be \[
	\wp(z) = \frac{1}{z^2} + \sum_{\substack{\omega \in \Lambda \\ \omega \ne 0}}\left( \frac{1}{(z-\omega)^2} - \frac{1}{\omega^2} \right) 
.\] 
At first glance it seems obvious that such a function is $\Lambda$ periodic, but we have to be careful. 
While this sum is absolutely convergent on compact sets, its is not if we split up $\frac{1}{(z-\omega^2)} - \frac{1}{\omega^2}$ into two different summations. So we cannot simply permute the terms in the desired way. 
It still is $\Lambda$-periodic. Indeed, notice that  \[
	\wp'(z) = -2 \sum_{\omega \in \Lambda} \frac{1}{(z - \omega)^3}
\] 
this is periodic function as we can permute the terms in the sum.
We will now show that this implies that $\wp$ is $\Lambda$-periodic (exercise 1.4.2).
Note that $\wp$ is symmetric, as showing this can indeed be achieved by permuting the right terms.
For any $\omega \in \Lambda$ is holds that 
\[
	(\wp(z) - \wp(z + \omega))' = \wp'(z) - \wp'(z+ \omega) = 0 
.\] 
From this it follows that $\wp(z) - \wp(z + \omega)$ is constant a constant function of $z$, everywhere where it is defined.
In particular it is the case that \[
	\wp\left(-\frac{\omega}{2}\right) - \wp\left(\omega - \frac{\omega}{2}\right) = \wp\left(\frac{\omega}{2}\right) - \wp\left(\frac{\omega}{2}\right)  = 0 
.\] 
So $\wp(z) = \wp(z+ \omega)$. 
We conclude that $\wp$ is $\Lambda$-periodic.

Implicitly $\wp$ depends on the lattice $\Lambda$. 
From now on we will write  $\wp_\Lambda$ to make it clear which lattice is used in the construction of $\wp$. 
Recall from \cref{sec:modular_forms_and_complex_tori} that every lattice is isomorphic to a lattice of the form $\Lambda_\tau = \tau\Z \oplus \Z$. For these lattices we will denote $\wp_{\Lambda_\tau}$ simply as $\wp_\tau$.

Recall that \emph{Eisenstein-series} $G_k$ is the function defined as \[
	G_k(\tau) = \sum_{\substack{c, d \in \Z^2 \\ (c, d) \ne 0}} \frac{1}{(c\tau + d)^{k}}
.\] 
Note that this is equal to \[
	G_k(\tau) = \sum_{\substack{\omega \in \Lambda_\tau \\ \omega \ne 0}} \frac{1}{\omega}
.\] 
In this spirit we extend the definition of the eisentein series as follows. For a given lattice $\Lambda$ and an even number  $k > 2$ we define \[
	G_k(\Lambda) = \sum_{\substack{\omega \in \Lambda \\ \omega \ne 0}} \frac{1}{\omega^{k}}
.\] 
Note that this is a very natural extension as $G_k(\tau) = G_k(\Lambda_\tau)$.
\section{Appendices}
\begin{theorem}
	The modular group $\SL_2(\Z)$ is generated by \[ 	
	\begin{pmatrix}  1 & 1 \\ 0 & 1 \end{pmatrix} 
	\text{ and }
	\begin{pmatrix} 0 & -1 \\ 1 & 0 \end{pmatrix} 
.\] 
\end{theorem}
\begin{proof}
	Let $\mathcal{A} $ be the group generated by the two matrices. 
	Suppose that $\SL_2(\Z) \setminus \mathcal{A} $ is not empty.
	Then we can choose a matrix \[
		A =\begin{pmatrix} a & b \\ c& d\end{pmatrix} \in \SL_2\Z \setminus \mathcal{A} 
	\] 
	with mimimal 1-norm, i.e. $|a| + |b| + |c| + |d|$ is minimal.
	We know that 
\end{proof}
\end{document}